Profit Optimisation in Heterogeneous 2-link Parallel Games

### 2025-01-08 thesis/article.tex
Thesis: Final fixes in Abstract, Introduction and Chapter 4
Release v1.0

\thispagestyle{empty}
\pagenumbering{gobble}
\chapter*{Abstract}
In this masters thesis we study a $2$-level game where, on an $n$-link network from a source $s$ to a target $t$, a unit of flow wants to move from $s$ to $t$ using the link with the lowest cost, while the link operators compete for profit by assigning tolls to links and thus creating a toll congestion game.
We examine only affine latencies for the links.
In addition, each flow user responds to tolls in a heterogeneous way, i.e. each player $p$ in the flow has a different time-money sensitivity value, which is described by a distribution function $\alpha(p)$.
We first present some definitions and properties of pricing homogeneous games, heavily based on previous work, and using them as base, we extend them to describe pricing games with heterogeneous users.
We introduce a new term, the sensitivity split function $\as(t)$, defined for any $2$ links of the network as the time-money sensitivity value for which, given a set of tolls $t$, the link costs are equal.
With the help of $\as$ we describe some properties around the selfish behaviour of heterogeneous users for different tolls, as well as prove the existence of bounds for the sensitivity values that are relevant to the game.
Then, arriving at the main body of this study, we investigate the existence of a Nash Equilibrium in the profit game between the link operators for different cases of heterogeneous users.
We show there is no Nash Equilibrium when there are many users with $0$ sensitivity, i.e. players not affected by tolls, so we further assume $\alpha(p) > 0$.
We then investigate cases of fixed distribution functions, which, even though create homogeneous behaviour for the users (making them pseudo-heterogeneous games), help us describe how flow, tolls and equilibria behave as that fixed value changes.
Using all the above, we study step distribution functions, where by limiting our scope to $2$ links, we present strict conditions necessary for the existence or not of a Nash Equilibrium in the profit game between the link operators.
We present and analyse an algorithm to calculate that equilibrium, if it exists, and finish with some discussion about potential future work.
\clearpage

@@

\chapter*{Συνοψη}
Σε αυτή τη διπλωματική μεταπτυχιακού μελετάμε ένα παίγνιο $2$ επιπέδων όπου, σε ένα δίκτυο με $n$ ακμές από μια πηγή $s$ σε ένα στόχο $t$, μια μονάδα ροής θέλει να μετακινηθεί από το $s$ στο $t$ χρησιμοποιώντας την ακμή με το χαμηλότερο κόστος, ενώ οι διαχειριστές των ακμών ανταγωνίζονται για κέρδος βάζοντας δίοδια στις ακμές και δημιουργώντας έτσι ένα παίγνιο συμφόρησης με διόδια.
Εξετάζουμε μόνο αφινικές καθυστερήσεις για τις ακμές.
Επιπλέον, ο κάθε χρήστης δικτύου αντιδρά στα διόδια με ετερογενή τρόπο, δηλ. ο κάθε παίκτης $p$ στη ροή του δικτύου έχει διαφορετική τιμή ευαισθησίας χρόνου-χρήματος, η οποία περιγράφεται από μια συνάρτηση κατανομής $\alpha(p)$.
Πρώτα παρουσιάζουμε κάποιους βασικούς ορισμούς και ιδιότητες των ομογενών παιγνίων κέρδους, κυρίως βασισμένα σε προηγούμενες δουλειές, και με αυτά ως βάση, τα επεκτείνουμε για να περιγράψουμε παίγνια κέρδους με ετερογενείς χρήστες.
Εισαγάγουμε έναν νέο όρο, τη συνάρτηση διαχωρισμού ευαισθησίας $\as(t)$, ορισμένη για οποιεσδήποτε $2$ ακμές του δικτύου ως η τιμή ευαισθησίας χρόνου-χρήματος για την οποία, δωσμένων σετ διοδίων $t$, τα κόστη των $2$ ακμών είναι ίσα.
Με τη βοήθεια της $\as$ θα περιγράψουμε μερικές ιδιότητες γύρω από την εγωιστική συμπεριφορά των ετερογενών χρηστών για διαφορετικά διόδια, καθώς και θα αποδείξουμε την ύπαρξη άνω φράγματος για τις τιμές ευαισθησίας που αφορούν το παίγνιο.
Στη συνέχεια, φτάνοντας στο κύριο σώμα αυτής της έρευνας, εξετάζουμε την ύπαρξη ισορροπίας Nash στο παίγνιο κέρδους μεταξύ των διαχειριστών των ακμών για διαφορετικές περιπτώσεις ετερογενών χρηστών.
Δείχνουμε ότι δεν υπάρχει ισορροπία Nash όταν υπάρχουν πολλοί χρήστες με $0$ ευαισθησία, δηλ. παίκτες που δεν επηρρεάζονται από τα διόδια, οπότε στο εξής υποθέτουμε ότι $\alpha(p) > 0$.
Κατόπιν εξετάζουμε περιπτώσεις όπου η συνάρτηση κατανομής είναι σταθερή, η οποίες, αν και δημιουργούν ομογενή συμπεριφορά στους χρήστες (κάνοντάς τα ψευδο-ετερογενή παίγνια), μας βοηθάνε να περιγράψουμε πώς συμπεριφέρονται η ροή, τα διόδια και τα σημεία ισορροπίας όταν αυτή η σταθερή τιμή μεταβάλλεται.
Χρησιμοποιώντας όλα τα παραπάνω, μελετάμε βηματικές συναρτήσεις κατανομής, όπου περιορίζοντας το μοντέλο μας σε στιγμιότυπα $2$ ακμών, παρουσιάζουμε αυστηρές συνθήκες αναγκαίες για την ύπαρξη ή μη σημείου ισορροπίας Nash στο παίγνιο κέρδους μεταξύ των διαχειριστών των ακμών.
Παρουσιάζουμε και αναλύουμε έναν αλγόριθμο για τον υπολογισμό του σημείου ισορροπίας, αν υπάρχει, και κλείνουμε με μια συζήτηση για πιθανή μελοντική δουλειά.

@@

In this thesis we study a two level game.
On the basis there is a network with 2 nodes $(s, t)$ and $n$ links, each link directed from $s$ to $t$ and weighted by a non-decreasing latency function that defines how the traffic increases as more players flow into that link.
On the lower level, a flow $[0, 1]$ wants to move from $s$ to $t$ using one of the $n$ links and experience minimum latency (traffic).
On the upper level, each link is managed by a different operator who can assign a toll to their link and gain profit proportional to that toll and the flow that ends up using the link.
The tolls impose an additional cost to the flow players, affecting the lower level game, and the link operators compete with each other to maximise their profit.
In our case we will additionally assume that the flow users behave in a heterogeneous way when it comes to tolls, meaning that each player has a different time-money trade-off and thus see the link costs differently.

@@

Entering the main body of our work, in Chapter \ref{chapter:split} we introduce a new term regarding any two links of the network, called \textit{money sensitivity split}, which captures the notion of a (potentially non-existent) player whose money sensitivity makes them view the two links as equal.
We will then analyse that notion and provide properties and bounds for its values.
Finally, in the chapter's main result, we prove a general property of heterogeneous network games as, due to the previously found bounds, there might be a portion of users whose sensitivity value can be arbitrarily large without affecting the game.

Having arrived at our main results in this work, in Chapter \ref{chapter:pricing_equilibria} we investigate the pricing competition game between the link operators of a network with heterogeneous users.
We begin by examining a class of games where the distribution function is fixed across all users, which, as we'll formally describe, makes the game \textit{pseudo-heterogeneous}.
Particular interest in the above class is the relation between games with identical setup but with different fixed distribution functions.
We describe and prove some properties for them, and using those we then tackle step distribution functions (where the user sensitivity can take values out of a finite set).
At this point we restrict our model to instances with $2$ links, and provide strict restrictions for equilibrium existence for the pricing competition game.

We conclude this thesis with a short discussion around the main outtakes of this study, while also presenting potential directions for future work.

@@

One other difficulty we need to consider are cases where $t_i = t_j$ for some links $i, j \in N$.
Remember that $x(t)$ is only uniquely defined with regards to costs, which in extension makes players in such links with equal cost (viewed as such by all players of the game) able to be put arbitrarily in any of the two, regardless of their $\alpha$ value.
Even if we attempted to assume a link assignment ordered by $\alpha$, the link order also matters, as the (equal latency) links might have different flow volume, resulting in different $\alpha$ values in the flow that uses each.
We tackle this by assuming that our sets contain all possible $\alpha$ values of any player that might use that link.
We therefore ignore the information about the actual flow splits among those links and keep only the ones among links with different toll values (and in consequence different latencies, Lemma \ref{lemma:latencies_tolls}).
%Still, even with this information loss, we observe that
%\begin{enumerate}
%	\item The notion of $\as$ as the $\alpha$ value that views two links as equal is still valid, since such a player will also view as equal any other link with the same toll value as any of the initial two.
%	\item If we limit our scope to only include the links with equal tolls, we actually get a homogeneous sub-game for the total flow in those links.
%\end{enumerate}

@@

All the above lemmas can now help us describe the behaviour of $\as$ for different tolls, and more specifically how it behaves with regards to $t_i$ for a fixed $t_{-i}$.
The main challenge in this analysis is that a changing $t_i$ can affect the sensitivity split of multiple link pairs, while even the link ordering is not always fixed, as we can put link $i$ at any place by selecting an appropriate $t_i$.
In order to help with the intuition, we will simplify our final analysis to $2$-link networks, where we essentially have two possible toll orderings, thus giving us a simplified look on what happens when that ordering changes.
Note that the concepts presented here are still valid for $n$-link networks, if we simply focus on consecutive links.
We start with the following corollary.

@@

In this chapter we will introduce the pricing competition game to our model and examine how different distribution functions can affect the profit maximisation game between the toll operators.
We will begin by noticing that players with no toll fear, i.e. $\alpha(p)$ close to $0$, can be exploited by some link operators who then get unbounded profits (Lemma \ref{lemma:a_0_0}).
This obviously breaks the pricing competition game and, in turn, leads us to include the restriction of $\alpha(p) > 0$ when $p > 0$.
We will then proceed to analyse games with fixed distribution functions, where in Lemma \ref{lemma:a_fixed_homogeneous} we will show that such games are, as intuition suggests, equivalent to homogeneous games.
Special interest will then be on comparing such games, specifically those with the same setup but different fixed $\alpha$ functions, as Lemma \ref{lemma:a_fixed_c_relation} will reveal that, uniformly varying the player sensitivity, maps the toll values around the same flow.
The tools and notions built above will prove extremely useful when we next analyse games with step distribution functions, where we'll restrict our analysis to instances with $2$ links to simplify our work.
For those we will show an equivalence with a homogeneous game for a given toll \ref{lemma:heterogenous_flow_equiv_fixed} and combining it with Lemma \ref{lemma:alpha_step_gaps} where we'll show that a Nash Equilibrium cannot exist with all players of each sensitivity type using a single link, we get this chapter's main result, Theorem \ref{theorem:a_step_ne_equiv_fixed}, which states that if a Nash Equilibrium exists in a heterogeneous game with a step distribution function, then the respective homogeneous also has an equilibrium for the same set of tolls.
Using this notion, we will describe an algorithm (\ref{algorithm:step_ne_pricing}) which can decide existence and return the Nash Equilibrium for the pricing competition game in a heterogeneous $2$-link network.

@@

\begin{proof}
	It suffices to show that $x^{(1)}(t) = x^{(2)}(t), \; \forall t \in \R_+^{|N|}$, where $x^{(1)}(t), x^{(2)}(t)$ the respective Nash Equilibrium for $t$ in $\Gm_1, \Gm_2$.
	Since $\Gm_2$ is homogeneous, we can use Lemma \ref{lemma:wardrop_equilibrium} to get that for any other flow $x^{(2)}$ it holds that
	\[
		\sum_{i \in N} \left(\tfrac1c \cdot \ell_i^{(1)}(x_i^{(2)}(t)) + t_i\right) \cdot \left(x_i^{(2)}(t) - x_i^{(2)}\right) \leq 0
	\]
	which is equivalent to
	\[
		\sum_{i \in N} \left(\ell_i^{(1)}(x_i^{(2)}(t)) + c t_i\right) \cdot \left(x_i^{(2)}(t) - x_i^{(2)}\right) \leq 0
	\]
	Notice that $\ell_i^{(1)}(x_i) + c t_i$ is the link cost of link $i$ for players in $\Gm_1$, given flow $x_i$ and toll $t_i$.
	Regardless of the heterogeneity of our users, since the above inequality holds, then for flow $x^{(2)}(t)$ in $\Gm_1$ it will also hold for any other flow $x_i^{(1)} = x^{(2)}$.
	It follows that any flow movement from $x^{(2)}(t)$ will again put players in links with larger link costs, making $x^{(2)}(t)$ a Nash Equilibrium for $t$ in $\Gm_1$ as well.
	Uniqueness of the equilibrium with regards to costs finally gives us $x^{(1)}(t) = x^{(2)}(t)$.
	Since the above hold for all $t \in \R_+^{|N|}$, we conclude from Definition \ref{definition:game_equivalence} that $\Gm_1, \Gm_2$ are equivalent.
\end{proof}
Since a heterogeneous game with a fixed distribution function is, in practice, homogeneous with regard to the flow behaviour, we will also use, when relevant, the term \textit{pseudo-heterogeneous game} to describe such flows more easily.

@@

\begin{proof}
	First we verify that $t^* = \left(\tfrac59, \tfrac49\right)$ is indeed a valid candidate pricing Nash equilibrium.
	Theorem \ref{theorem:a_step_ne_equiv_fixed} requires $t^*$ to also be a Nash Equilibrium at the pricing game of $\Gm^c = ([2], \ell, c)$, a pseudo-heterogeneous game where $c = 2$ or $3$.
	Since Lemma \ref{lemma:a_fixed_pricing_ne_relation} suggests that the flow at that equilibrium will be the same regardless of the value of $c$, we calculate the one from the homogeneous game $G^{\prime} = ([2], \ell, 1)$.
	We can calculate the latter equilibrium $t^{*\prime}$ by solving the system of equations with \ref{eq:homogeneous_x_i} and \ref{eq:homogeneous_pricing_ne_i}, from which we get $t^{*\prime} = \left(\tfrac53, \tfrac43\right)$ with $x^\prime(t^{*\prime}) = \left(\tfrac59, \tfrac49\right)$.
	Such a flow split puts the flow of link $1$ in the upper split and link $2$ in the lower split, making $\as(t^*) = \alpha(x_1(t^{*\prime})) = \alpha(5/9) = 4$.
	Lemma \ref{lemma:a_fixed_pricing_ne_relation} then gives us
	\[
		t^* = \frac13 t^{*\prime} = \left(\frac{5}{12}, \frac{4}{12}\right) \qquad x(t^*) = x^\prime(t^{*\prime}) = \left(\frac59, \frac49\right)
	\]
	We're only left to check whether $t^*$ satisfies the definition of a Nash Equilibrium, i.e. whether each link operator has a max profit toll in response to the opposing one.
	We will therefore need to calculate the profit functions for each link, for which, thanks to \ref{lemma:heterogenous_flow_equiv_fixed}$(ii)$, we can use the respective pseudo-heterogeneous profit functions for each $c = 2, 3$.
	We can then take the profit definition, formula \ref{eq:homogeneous_x_i} for $2$ links as well as Lemma \ref{lemma:a_fixed_c_relation}$(ii)$ and get for $\Gm^c$
	\[
		\Pi_1^c(t_1, t_2^*) = t_1 \cdot \frac{2 + c (4/3 - t_1)}{3}
		\qquad
		\Pi_2^c(t_1^*, t_2) = t_2 \cdot \frac{1 + c (5/3 - t_2)}{3}
	\]
	Next, we need to calculate the toll difference intervals for which $\as(t) = 2, 3$.
	Similarly to Example \ref{example:a_step_simple}, for each link operator we fix the opposing toll $t_{-i}^*$ and consider all flow cases where all players of each sensitivity type are using a single link, i.e. $x = \left(\tfrac14, \tfrac34\right)$ and $\left(\tfrac34, \tfrac14\right)$, for all possible $\alpha$ values, i.e. $2$ and $3$.
	We ignore the cases $x = (1, 0), (0, 1)$ as they minimise profit for both link operators.
	For toll differences where $\as(t) = c$, then profit is equal to $\Pi_i^c(t)$ (Lemma \ref{lemma:heterogenous_flow_equiv_fixed}).
	Otherwise, the flow remains static, making $x_i t_i$ a strictly increasing linear function with regards to $t_i$.

@@

\begin{figure}
	\centering
	\begin{tikzpicture}[x=10cm, y=17cm, every node/.style={font=\small}]
		% Axes
		\draw[->] (0,0) -- (-0.4,0);
		\draw[->] (0,0) -- (0.7,0) node[right] {$t_1 - t_2$};
		\draw[->] (0,0) -- (0,0.3) node[above] {$\Pi_1(t)$};

>>>>

\thispagestyle{empty}
\pagenumbering{gobble}
\chapter*{Abstract}
In this masters thesis we study a $2$-level game where, on an $n$-link network from a source $s$ to a target $t$, a unit of flow wants to move from $s$ to $t$ using the link with the lowest cost, while the link operators compete for profit by assigning tolls to links and thus creating a toll congestion game.
We examine only affine latencies for the links.
In addition, each flow user responds to tolls in a heterogeneous way, i.e. each player $p$ in the flow has a different time-money sensitivity value, which is described by a distribution function $\alpha(p)$.
We first present some definitions and properties of pricing homogeneous games, heavily based on previous work, and using them as base, we extend them to describe pricing games with heterogeneous users.
We introduce a new term, the sensitivity split function $\as(t)$, defined for any $2$ links of the network as the time-money sensitivity value for which, given a set of tolls $t$, the link costs are equal.
With the help of $\as$ we describe some properties around the selfish behaviour of heterogeneous users for different tolls, as well as prove the existence of bounds for the sensitivity values that are relevant to the game.
Then, arriving at the main body of this study, we investigate the existence of a Nash Equilibrium in the profit game between the link operators for different cases of heterogeneous users.
We show there is no Nash Equilibrium when there are many users with $0$ sensitivity, i.e. players not affected by tolls, forcing us to assume onward $\alpha(p) > 0$.
We then investigate cases of fixed distribution functions, which create homogeneous behaviours for the users (making them pseudo-heterogeneous games), through which we then describe how flow, tolls and equilibria behave as that fixed value changes.
Using all the above, we study step distribution functions, where, by limiting our scope to $2$ links, we prove a strong requirement for the value of a potential Nash Equilibrium in the profit game between the link operators, connecting it to the equilibrium of an equivalent pseudo-heterogeneous game.
We finally present and analyse an algorithm that calculates that equilibrium, if it exists, in time linear with regard to the step function's cardinality, and we finish with some discussion about potential future work.
\clearpage

@@

\chapter*{Συνοψη}
Σε αυτή τη διπλωματική μεταπτυχιακού μελετάμε ένα παίγνιο $2$ επιπέδων όπου, σε ένα δίκτυο με $n$ ακμές από μια πηγή $s$ σε ένα στόχο $t$, μια μονάδα ροής θέλει να μετακινηθεί από το $s$ στο $t$ χρησιμοποιώντας την ακμή με το χαμηλότερο κόστος, ενώ οι διαχειριστές των ακμών ανταγωνίζονται για κέρδος βάζοντας δίοδια στις ακμές και δημιουργώντας έτσι ένα παίγνιο συμφόρησης με διόδια.
Εξετάζουμε μόνο αφινικές καθυστερήσεις για τις ακμές.
Επιπλέον, ο κάθε χρήστης δικτύου αντιδρά στα διόδια με ετερογενή τρόπο, δηλ. ο κάθε παίκτης $p$ στη ροή του δικτύου έχει διαφορετική τιμή ευαισθησίας χρόνου-χρήματος, η οποία περιγράφεται από μια συνάρτηση κατανομής $\alpha(p)$.
Πρώτα παρουσιάζουμε κάποιους βασικούς ορισμούς και ιδιότητες των ομογενών παιγνίων κέρδους, κυρίως βασισμένα σε προηγούμενες δουλειές, και με αυτά ως βάση, τα επεκτείνουμε για να περιγράψουμε παίγνια κέρδους με ετερογενείς χρήστες.
Εισαγά-γουμε έναν νέο όρο, τη συνάρτηση διαχωρισμού ευαισθησίας $\as(t)$, ορισμένη για ο-ποιεσδήποτε $2$ ακμές του δικτύου ως η τιμή ευαισθησίας χρόνου-χρήματος για την οποία, δωσμένων σετ διοδίων $t$, τα κόστη των $2$ ακμών είναι ίσα.
Με τη βοήθεια της $\as$ θα περιγράψουμε μερικές ιδιότητες γύρω από την εγωιστική συμπεριφορά των ετερογενών χρηστών για διαφορετικά διόδια, καθώς και θα αποδείξουμε την ύπαρξη άνω φράγματος για τις τιμές ευαισθησίας που αφορούν το παίγνιο.
Στη συνέχεια, φτάνοντας στο κύριο σώμα αυτής της έρευνας, εξετάζουμε την ύπαρξη ισορροπίας Nash στο παίγνιο κέρδους μεταξύ των διαχειριστών των ακμών για διαφορετικές περι-πτώσεις ετερογενών χρηστών.
Δείχνουμε ότι δεν υπάρχει ισορροπία Nash όταν υπάρχουν πολλοί χρήστες με $0$ ευαισθησία, δηλ. παίκτες που δεν επηρρεάζονται από τα διόδια, το οποίο μας αναγκάζει στο εξής να υποθέσουμε ότι $\alpha(p) > 0$.
Κατόπιν, εξετάζουμε περιπτώσεις όπου η συνάρτηση κατανομής είναι σταθερή, η οποίες δημιουργούν ομογενή συμπεριφορά στους χρήστες (κάνοντάς τα ψευδο-ετερογενή παίγνια), μέσω των οποίων μετά θα περιγράψουμε πώς συμπεριφέρονται η ροή, τα διόδια και τα σημεία ισορροπίας όταν αυτή η σταθερή τιμή μεταβάλλεται.
Χρησιμοποιώντας όλα τα παραπάνω, μελετάμε βηματικές συναρτήσεις κατανομής, όπου, περιορίζοντας το μοντέλο μας σε στιγμιότυπα $2$ ακμών, αποδεικνύουμε μια αυστηρή συνθήκη για την τιμή ενός πιθανού σημείου ισορροπίας Nash στο παίγνιο κέρδους μεταξύ των διαχειριστών των ακμών, συνδέοντάς το με το σημείο ισορροπίας σε ένα ισοδύναμο ψευδο-ετερογενή παίγνιο.
Εντέλει, παρουσιάζουμε και αναλύουμε έναν αλγόριθμο για τον υπολογισμό αυτού του σημείου ισορροπίας, αν υπάρχει, σε χρόνο γραμμικό σε σχέση με τον πληθάριθμο της βηματικής συνάρτησης, και κλείνουμε με μια συζήτηση για πιθανές μελλοντικές εργασίες.

@@

In this thesis we study a two level game.
On the basis there is a network with 2 nodes $(s, t)$ and $n$ links, each link directed from $s$ to $t$ and weighted by a non-decreasing latency function that defines how the traffic increases as more players flow into that link.
On the lower level, a flow $[0, 1]$ wants to move from $s$ to $t$ using one of the $n$ links and experience minimum latency (traffic).
On the upper level, each link is managed by a different operator who can assign a toll to their link and gain profit proportional to that toll and the flow that ends up using the link.
The tolls impose an additional cost to the flow players, affecting the lower level game, and the link operators compete with each other to maximise their profit.
In our case, we will additionally assume that the flow users behave in a heterogeneous way when it comes to tolls, meaning that each player has a different time-money trade-off and thus see the link costs differently.

@@

Entering the main body of our work, in Chapter \ref{chapter:split} we introduce a new term regarding any two links of the network, called \textit{money sensitivity split}, which captures the notion of a (potentially non-existent) player whose money sensitivity makes them view the two link costs as equal.
We will then analyse that notion and provide properties and bounds for its values.
Finally, in the chapter's main result, we prove a general property of heterogeneous network games as, due to the previously found bounds, there might be a portion of users whose sensitivity value can be arbitrarily large without affecting the game.

Having arrived at our main results in this work, in Chapter \ref{chapter:pricing_equilibria} we investigate the pricing competition game between the link operators of a network with heterogeneous users.
We begin by examining a class of games where the distribution function is fixed across all users, which, as we'll formally describe, makes the game \textit{pseudo-heterogeneous}.
Particular interest in the above class is the relation between games with identical setup but with different fixed distribution functions.
We describe and prove some properties for them, and using those we then tackle step distribution functions (where the user sensitivity can take values out of a finite set).
At this point we restrict our model to instances with $2$ links, and prove a strong condition about the value of a potential Nash Equilibrium for the pricing game, connecting it to the respective equilibrium of an equivalent pseudo-heterogeneous game.
We will discuss with examples the consequences of such a condition, and finally, we will present, analyse and prove the validity of an algorithm which can decide on the existence as well as calculate a Nash Equilibrium for the pricing game in $\mathcal{O}(m)$ time (with $m$ being the size of the step distribution function's finite set of values).

We conclude this thesis with a short discussion around potential directions for future work.

@@

One other difficulty we need to consider are cases where $t_i = t_j$ for some links $i, j \in N$.
Remember that $x(t)$ is only uniquely defined with regard to costs, which in extension makes players in such links with equal cost (viewed as such by all players of the game) able to be put arbitrarily in any of the two, regardless of their $\alpha$ value.
Even if we attempted to assume a link assignment ordered by $\alpha$, the link order also matters, as the (equal latency) links might have different flow volume, resulting in different $\alpha$ values in the flow that uses each.
We tackle this by assuming that our sets contain all possible $\alpha$ values of any player that might use that link.
We therefore ignore the information about the actual flow splits among those links and keep only the ones among links with different toll values (and in consequence different latencies, Lemma \ref{lemma:latencies_tolls}).
%Still, even with this information loss, we observe that
%\begin{enumerate}
%	\item The notion of $\as$ as the $\alpha$ value that views two links as equal is still valid, since such a player will also view as equal any other link with the same toll value as any of the initial two.
%	\item If we limit our scope to only include the links with equal tolls, we actually get a homogeneous sub-game for the total flow in those links.
%\end{enumerate}

@@

All the above lemmas can now help us describe the behaviour of $\as$ for different tolls, and more specifically how it behaves with regard to $t_i$ for a fixed $t_{-i}$.
The main challenge in this analysis is that a changing $t_i$ can affect the sensitivity split of multiple link pairs, while even the link ordering is not always fixed, as we can put link $i$ at any place by selecting an appropriate $t_i$.
In order to help with the intuition, we will simplify our final analysis to $2$-link networks, where we essentially have two possible toll orderings, thus giving us a simplified look on what happens when that ordering changes.
Note that the concepts presented here are still valid for $n$-link networks, if we simply focus on consecutive links.
We start with the following corollary.

@@

In this chapter we will introduce the pricing competition game to our model and examine how different distribution functions can affect the profit maximisation game between the link operators.
We will begin by noticing that players with no toll fear, i.e. $\alpha(p)$ close to $0$, can be exploited by some link operators who then get unbounded profits (Lemma \ref{lemma:a_0_0}).
This obviously breaks the pricing competition game and, in turn, leads us to include the restriction of $\alpha(p) > 0$ when $p > 0$.
We will then proceed to analyse games with fixed distribution functions, where in Lemma \ref{lemma:a_fixed_homogeneous} we will show that such games are, as intuition suggests, equivalent to homogeneous games.
Special interest will then be on comparing such games, specifically those with the same setup but different fixed $\alpha$ functions, as Lemma \ref{lemma:a_fixed_c_relation} will reveal that, uniformly varying the player sensitivity, maps the toll values around the same flow.
The tools and notions built above will prove extremely useful when we next analyse games with step distribution functions, where we'll restrict our analysis to instances with $2$ links to simplify our work.
We will show that $2$-link games, for a fixed set of tolls, are equivalent to a specific pseudo-heterogeneous game  (Lemma \ref{lemma:heterogenous_flow_equiv_fixed}) and combining this property with Lemma \ref{lemma:alpha_step_gaps}, where we'll show that a Nash Equilibrium cannot exist with all players of each sensitivity type using a single link, we get this chapter's main result, Theorem \ref{theorem:a_step_ne_equiv_fixed}, which states that if a Nash Equilibrium exists in a heterogeneous game with a step distribution function, then a respective pseudo-heterogeneous game also has an equilibrium for the same set of tolls.
Using this notion, we will describe an algorithm (\ref{algorithm:step_ne_pricing}) which can decide existence and return the Nash Equilibrium for a $2$-link pricing competition game in $\mathcal{O}(m)$ time.

@@

\begin{proof}
	It suffices to show that $x^{(1)}(t) = x^{(2)}(t), \; \forall t \in \R_+^{|N|}$, where $x^{(1)}(t), x^{(2)}(t)$ the respective Nash Equilibrium for $t$ in $\Gm_1, \Gm_2$.
	Since $\Gm_2$ is homogeneous, we can use Lemma \ref{lemma:wardrop_equilibrium} to get that for any other flow $x^{(2)}$ it holds that
	\[
		\sum_{i \in N} \left(\tfrac1c \cdot \ell_i^{(1)}(x_i^{(2)}(t)) + t_i\right) \cdot \left(x_i^{(2)}(t) - x_i^{(2)}\right) \leq 0
	\]
	which is equivalent to
	\[
		\sum_{i \in N} \left(\ell_i^{(1)}(x_i^{(2)}(t)) + c t_i\right) \cdot \left(x_i^{(2)}(t) - x_i^{(2)}\right) \leq 0
	\]
	Notice that $\ell_i^{(1)}(x_i) + c t_i$ is the link cost of link $i$ for players in $\Gm_1$, given flow $x_i$ and toll $t_i$.
	Regardless of the heterogeneity of our users, since the above inequality holds, then for flow $x^{(2)}(t)$ in $\Gm_1$ it will also hold for any other flow $x_i^{(1)} = x^{(2)}$.
	It follows that any flow movement from $x^{(2)}(t)$ will again put players in links with larger link costs, making $x^{(2)}(t)$ a Nash Equilibrium for $t$ in $\Gm_1$ as well.
	Uniqueness of the equilibrium with regard to costs finally gives us $x^{(1)}(t) = x^{(2)}(t)$.
	Since the above hold for all $t \in \R_+^{|N|}$, we conclude from Definition \ref{definition:game_equivalence} that $\Gm_1, \Gm_2$ are equivalent.
\end{proof}
Since a heterogeneous game with a fixed distribution function is, in practice, homogeneous with regard to the flow behaviour, we will also use, when relevant, the term \textit{pseudo-heterogeneous game} to describe such flows more easily.

@@

\begin{proof}
	First we verify that $t^* = \left(\tfrac{5}{12}, \tfrac{4}{12}\right)$ is indeed a valid candidate pricing Nash equilibrium.
	Theorem \ref{theorem:a_step_ne_equiv_fixed} requires $t^*$ to also be a Nash Equilibrium at the pricing game of $\Gm_c = ([2], \ell, c)$, a pseudo-heterogeneous game where $c = 2$ or $3$.
	Since Lemma \ref{lemma:a_fixed_pricing_ne_relation} suggests that the flow at that equilibrium will be the same regardless of the value of $c$, we calculate the one from the homogeneous game $G^{\prime} = ([2], \ell, 1)$.
	We can calculate the latter equilibrium $t^{*\prime}$ by solving the system of equations with \ref{eq:homogeneous_x_i} and \ref{eq:homogeneous_pricing_ne_i}, from which we get $t^{*\prime} = \left(\tfrac53, \tfrac43\right)$ with $x^\prime(t^{*\prime}) = \left(\tfrac59, \tfrac49\right)$.
	Such a flow split puts the flow of link $1$ in the lower split and link $2$ in the upper split, making $\as(t^*) = \alpha(x_1(t^{*\prime})) = \alpha(5/9) = 4$.
	Lemma \ref{lemma:a_fixed_pricing_ne_relation} then gives us
	\[
		t^* = \frac13 t^{*\prime} = \left(\frac{5}{12}, \frac{4}{12}\right) \qquad x(t^*) = x^\prime(t^{*\prime}) = \left(\frac59, \frac49\right)
	\]
	We're only left to check whether $t^*$ satisfies the definition of a Nash Equilibrium, i.e. whether each link operator has a max profit toll in response to the opposing one.
	We will therefore need to calculate the profit functions for each link, for which, thanks to \ref{lemma:heterogenous_flow_equiv_fixed}$(ii)$, we can use the respective pseudo-heterogeneous profit functions for each $c = 2, 3$.
	We can then take the profit definition, formula \ref{eq:homogeneous_x_i} for $2$ links as well as Lemma \ref{lemma:a_fixed_c_relation}$(ii)$ and get for $\Gm_c$
	\[
		\Pi_{1, c}(t_1, t_2^*) = t_1 \cdot \frac{2 + c (4/3 - t_1)}{3}
		\qquad
		\Pi_{2, c}(t_1^*, t_2) = t_2 \cdot \frac{1 + c (5/3 - t_2)}{3}
	\]
	Next, we need to calculate the toll difference intervals for which $\as(t) = 2, 3$.
	Similarly to Example \ref{example:a_step_simple}, for each link operator we fix the opposing toll $t_{-i}^*$ and consider all flow cases where all players of each sensitivity type are using a single link, i.e. $x = \left(\tfrac14, \tfrac34\right)$ and $\left(\tfrac34, \tfrac14\right)$, for all possible $\alpha$ values, i.e. $2$ and $3$.
	We ignore the cases $x = (1, 0), (0, 1)$ as they minimise profit for both link operators.
	For toll differences where $\as(t) = c$, profit is equal to $\Pi_{i, c}(t)$ (Lemma \ref{lemma:heterogenous_flow_equiv_fixed}).
	Otherwise, the flow remains static, making $x_i t_i$ a linear function with regard to $t_i$, with slope $x_i > 0$.

@@

\begin{figure}
	\centering
	\begin{tikzpicture}[x=8cm, y=10cm, every node/.style={font=\small}]
		% Axes
		\draw[->] (0,0) -- (-0.4,0);
		\draw[->] (0,0) -- (0.7,0) node[right] {$t_1 - t_2$};
		\draw[->] (0,0) -- (0,0.3) node[above] {$\Pi_1(t)$};

		% Tick marks
		\draw (-1/3,0) -- (-1/3,-0.01) node[below] {$-\frac13$};
		\draw (-1/4,0) -- (-1/4,-0.01) node[below] {$-\frac14$};
		\draw (-1/12,0) -- (-1/12,-0.01) node[left=9,below] {$-\frac{1}{12}$};
		\draw (-1/16,0) -- (-1/16,-0.01) node[right=2,below] {$-\frac{1}{16}$};
		\draw (0,0) -- (0,-0.01) node[below] {$0$} node[below=20] {$t_2^* = \frac{4}{12}$};
		\draw (1/12,0) -- (1/12,-0.01) node[below] {$\frac{1}{12}$};
		\draw (5/16,0) -- (5/16,-0.01) node[below] {$\frac{5}{16}$};
		\draw (5/12,0) -- (5/12,-0.01) node[below] {$\frac{5}{12}$};
		\draw (2/4,0) -- (2/4,-0.01) node[below] {$\frac24$};
		\draw (2/3,0) -- (2/3,-0.01) node[below] {$\frac23$};
		%\draw (0,2/9) -- (-0.02,2/9) node[left] {$\frac29$};
		%\draw (0,1/12) -- (0.01,1/12) node[right] {$\frac{1}{12}$};

		% Pseudo-heterogeneous profits
		\draw[gray,dashed,thick,domain=-1/3:2/3] plot (\x,{((2-3*\x)/3)*(1/3+\x)}) node[above=17,right=-7,black] {$c = 3$};
		\draw[gray,dashed,thick,domain=-1/4:2/4] plot (\x,{((2-4*\x)/3)*(1/3+\x)}) node[above=10,right=-3,black] {$c = 4$};
